InnoDB Full text Search hacking with Shiny/R

### 2015-12-28 article.rmd
publishing for testing

---
title: "InnoDB Full text Search hacking with Shiny/R"
author: "Emanuel Calvo"
company: "Pythian"
date: "December 22, 2015"
output: html_document
runtime: shiny
---

```{r init, message = FALSE, echo=FALSE}
#options(width = 9999)
#knitr::opts_chunk$set(width=9999)

@@

@@

@@

## Moto of the article

@@

## Stats

@@

                DataResults = reactive({
                   f<-function(x) {as.character(input$filtering)}
                  sql <- paste("SELECT * FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") LIMIT 10", sep = "")

@@

## Recommendations

- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
  size is crucial when using this.
- FTS is not the best fit for exact string matches. You won't use FTS for searching emails or strings as words using
  single fields. 
- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the 
  FTS functions to avoid double executions. You can do this using somehting like:
  `match(content,title) against ("first (<second >third)") > 1 `
  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches. 
- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
  maintance on the table.
- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.
- If you want to have the results ordered by the last FTS expression in the WHERE clause, don't use `ORDER BY`, as MySQL
  forcely does an order internally. 
- Try to don't use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
  it consumes a considerably amount of CPU.

---------------------- FROM HERE ONLY TRASH TO BE REMOVED -----------------------------

Template

```{r}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),

  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")

  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package="shiny"),
  options=list(
    width="100%", height=550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.

@@

>>>>

---
title: "InnoDB Full text Search hacking with Shiny/R"
author: "Emanuel Calvo - Pythian"
date: "December 22, 2015"
output: html_document
runtime: shiny
---

```{r init, message = FALSE, echo=FALSE}
#options(width = 9999)
#knitr::opts_chunk$set(width=9999)
# opts_chunk$set(fig.width=9999) #, fig.height=6))

@@

## Version

 **BETA - NOT FINAL**

@@

Repository available at [Github](https://github.com/3manuek/fts_article).

## Credits

Author: Emanuel Calvo
Company: Pythian

Thanks to Valerie Parham-Thompson @ Pythian and Daniel Prince @ Oracle.

## Requirements

@@

## Some initial thoughts

@@

## Considerations if you are planning to use Innodb's Full text search

- Use QUERY EXPANSION only if you are interested to search relations over exact matches. Remember that the field
  size is crucial when using this.
- FTS is not the best fit for exact string matches. You won't use FTS for searching emails or strings as words using
  single fields. 
- If you are using BOOLEAN MODE, you can use the rank score to filter rows. MySQL is clever enough to optimize the 
  FTS functions to avoid double executions. You can do this using somehting like:
  `match(content,title) against ("first (<second >third)") > 1 `
  Generally, scores lower than 1 can be ignored when using boolean or natural mode searches. 
- `OPTIMIZE TABLE` does a rebuild of the table. To avoid this, set `innodb_optimize_fulltext_only=1` in order to do an incremental
  maintance on the table.
- Recall that NATURAL LANGUAGE MODE does not take the operands as the BOOLEAN MODE.
- If you plan to order by rank, it is not necessary to specify the clause `ORDER BY` as MySQL does the order forcelly internally. Also,
  the behavior is different from the default as it returns the heaviest at the top (like ORDER BY rank DESC).
- If you come from MyISAM's FTS implementation, recall that the ranking scoring is different.
- Try to don't use more than one FTS expression in the where clause. Keep in mind that this affects the order in the results and
  it consumes a considerably amount of CPU. MySQL orders by the latest expression in the WHERE clause. [WL#7123](https://dev.mysql.com/worklog/task/?id=7123)
- Also, if avoiding the rank information in the projection (SELECT clause) and using other aggregations like count(*), will use
  the "no ranking" FT_hints. The `limit` hint won't be used if invoked explicitely an `ORDER BY` and th `MATCH` clause in the projection.

```
(test) >  explain  select  * 
                      from bookContentByLine 
                      where match(content) against ("+home" IN BOOLEAN MODE) 
                      ORDER BY FTS_DOC_ID  
                      LIMIT 10\G
  select_type: SIMPLE
        table: bookContentByLine
         type: fulltext
        Extra: Using where; Ft_hints: no_ranking; Using filesort

(test) >  explain  select  * 
                      from bookContentByLine 
                      where match(content) against ("+home" IN BOOLEAN MODE) 
                      LIMIT 10\G
        table: bookContentByLine
         type: fulltext
        Extra: Using where; Ft_hints: no_ranking, limit = 10

(test) >  explain  select  count(content) 
                    from bookContentByLine 
                    where match(content) against ("+home" IN BOOLEAN MODE) \G
        table: bookContentByLine
         type: fulltext
        Extra: Using where; Ft_hints: no_ranking

```
- If you plan to use FTS_DOC_ID column with AUTO_INCREMENT option, have in mind that there is a limitation regarding this.You must
  declare a single column PRIMARY KEY constraint or as an UNIQUE index. Also, the data type is stricted as `bigint unsigned`. I.e:

```
CREATE TABLE test ( FTS_DOC_ID bigint unsigned  AUTO_INCREMENT, 
                    mainPk bigint, other text, 
                    PRIMARY KEY(mainPk),
                    UNIQUE(FTS_DOC_ID)
                    );
```

## Stats of the datasets

Here we prompt some stats regarding the used datasets for the tests and data written here:

@@

                DataResults = reactive({
                   f<-function(x) {as.character(input$filtering)}
                  sql <- paste("SELECT * FROM bookContentByLine WHERE MATCH(content) AGAINST(\"" ,f(x)," \" ",input$ftsMode,") ", sep = "")

                  rs <- head(dbGetQuery(con, sql), 50)
                  return(rs)

@@

@@

### Features introduces
